---
title: 'EA20190828 Report'
author: "Robin B. Trayler"
date: "8/29/2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(magrittr)
```
The $\delta^{13}C$ and $\delta^{15}N$ values of all samples were measured using a __Costech 4010__ Elemental Analyzer coupled with a __Delta V Plus Continuous Flow Isotope Ratio Mass Spectrometer__ housed in the Stable Isotope Facility at the University of California, Merced. If you would like the original Isodat 3.0 output file please request by sending an email to StableIsotopes@ucmerced.edu

# Run Parameters
```{r define_parameters, echo = TRUE}
########################################################################
file_path = "~/Desktop/EA/EA20190828.csv" # location of raw data file ##
########################################################################

file_name = tools::file_path_sans_ext(basename(file_path))
nitrogen_peak = 3 # which peak is nitrogen
carbon_peak = 4 # which peak is carbon
total_peaks = 6 # total number of peaks
linearity_standard = "USGS 40"
standards_used = c("USGS 41a", 'USGS 40') # standards used in run
reference_material = c("costech acetanilide") # material used for wt% reference
linearity_correction = FALSE # should a linearity correction be performed
linearity_standard = "USGS 40" # standard used for linearity correction
```

```{r, echo = FALSE}
# define values for standards
standards = tibble(sample = c('USGS 40', 'USGS 41a', 'costech acetanilide', 'peach', 'Em soil'),
                   d13C_true = c(-26.39, 36.55, -27.86, -25.99, -27.6),
                   d15N_true = c(-4.52, 47.55, -0.75, 1.98, -1.7),
                   C_wt_percent = c(40.8, 40.8, 71.09, 46, 8.68),
                   N_wt_percent = c(9.52, 9.52, 10.36, 2.82, 0.38))
```

```{r load_data, echo=FALSE, message=FALSE, warning=FALSE}
# Load data
data = read_csv(file = file_path)
# filter out the useful columns
data %<>% select(Identifier_1 = `Identifier 1`,
                 sample = `Identifier 2`,
                 wt_percent = `Amt%`,
                 amount = Amount,
                 `Area 28` = `Area 28`,
                 `Area 44` = `Area 44`,
                 d15N_measured = `d 15N/14N`,
                 d13C_measured = `d 13C/12C`,
                 `Peak Nr` = `Peak Nr.`)


```

# Check For Bad Samples
```{r check_data, echo = FALSE}
# Check for missing peaks
ID = unique(data[['Identifier_1']])

bad_peaks = rep(FALSE, length = length(ID))
for(i in 1:length(ID)) {
  if (data %>% filter(Identifier_1 == ID[i]) %>% select(`Peak Nr`) %>% nrow() < total_peaks) {
    bad_peaks[i] = TRUE
  }
}

```
Sample(s) with an Identifier 1 of __`r ID[bad_peaks]`__ have fewer peaks than expected. Please check the raw data. 


```{r select_standards,echo=FALSE}
# build a string of standard values for use in filtering the data later
# standard_string = character()
# for (i in 1:length(standards_used)) {
#   if (i >= 1 & i < length(standards_used)) {
#     standard_string = paste(standard_string, 'sample == ', standards_used[i], '|')
#   }
#   else
#     if (i == length(standards_used)) {
#       standard_string = paste(standard_string, 'sample == ', standards_used[i])
#     }
# }
# # add the reference material to the string
# reference_string = paste(standard_string,'| sample ==', reference_material)
```

```{r, echo=FALSE}
# filter and select
#Filter out the nitrogen and carbon data seperately. This is based on peak number which can cause problems if there are  #missing peaks. It might be better to filter on retention time so the code will only grab peaks within the correct time #window.
nitrogen = data %>% 
  filter(`Peak Nr` == nitrogen_peak) %>% 
  select(Identifier_1, 
         sample,
         N_wt_percent = wt_percent,
         amount,
         `Area 28`,
         d15N_measured,
         `Peak Nr`)

carbon = data %>% 
  filter(`Peak Nr` == carbon_peak) %>% 
  select(Identifier_1, 
         sample,
         C_wt_percent = wt_percent,
         amount,
         `Area 44`,
         d13C_measured,
         `Peak Nr`)
```

<!-- #Recalculate N and C percent -->
```{r include=FALSE}
#Recalculate N and C percent
# if (recalculate_wt_percent == TRUE) {
#   
#   C_wt_ref = as.numeric(standards %>% 
#                           filter(sample == 'USGS 40') %>% 
#                           select(C_wt_percent))
#   
#     C_wt_fit = carbon %>% 
#     filter(str_detect(sample, 'USGS 40')) %>% 
#     lm(amount * C_wt_ref/100  ~ `Area 44`, data = .)
#   
#   carbon %>% 
#     filter(str_detect(sample, 'USGS 40')) %>% 
#     with(plot(C_wt$`Area 44`, 
#          C_wt$amount * (C_wt_ref/100), 
#          xlab = 'Area 44',
#          ylab = 'carbon weight %',
#          tcl = 0.25,
#          type = 'p',
#          cex = 1.5,
#          pch = 21,
#          bg = 'steelblue'))
#   grid()
#   abline(C_wt_fit, lwd = 2)
#   C_wt_fit %<>% coef
#   carbon %<>% mutate(C_wt2 = (C_wt_fit[2] * `Area 44`  + C_wt_fit[1])/ amount * 100)
# }
```


<!-- # Linearity correction -->
<!-- Explination and plots to come soon. -->

```{r N_linearity, include=FALSE}
if (linearity_correction == TRUE) {
  # correct for linearity in nitrogen --------------------------------------------------
  # find samples with 'lin' prefix
  N_linearity <- nitrogen %>% filter(str_detect(sample, linearity_standard))
  # calculate coefficients
  N_coeff = N_linearity %>% lm(d15N_measured ~ `Area 28`, data = .) %>% coef()
  # calculate R squared
  N_summary = N_linearity %>% lm(d15N_measured ~ `Area 28`, data = .) %>% summary()
  # perform linearity correction
  nitrogen %<>% mutate(d15N_measured = 
                        N_coeff[1] + `Area 28` * N_coeff[2] -
                        as.numeric(standards %>% 
                                     filter(sample == linearity_standard) %>% 
                                     select(d15N_true)) + d15N_measured)
}
```

```{r C_linearity, include=FALSE}
if (linearity_correction == TRUE) {
  # correct for linearity in nitrogen --------------------------------------------------
  # find samples with 'lin' prefix
  C_linearity <- carbon %>% filter(str_detect(sample, linearity_standard))
  # calculate coefficients
  C_coeff = C_linearity %>% lm(d13C_measured ~ `Area 44`, data = .) %>% coef()
  # calculate R squared
  C_summary = C_linearity %>% lm(d13C_measured ~ `Area 44`, data = .) %>% summary()
  # perform linearity correction
  carbon %<>% mutate(d13C_measured = 
                      C_coeff[1] + `Area 44` * C_coeff[2] -
                      as.numeric(standards %>% 
                                   filter(sample == linearity_standard) %>% 
                                   select(d13C_true)) + d13C_measured)
}
```

# Standard correction
The isotope composition of unknowns samples were corrected to the international PDB ($\delta^{13}C$) and Air ($\delta^{15}N$) scales using the __`r paste(standards_used, collapse = ' and ')`__ standard reference materials. Total carbon and nitrogen contents were calculated using an in house reference material, __`r paste(reference_material)`__ which was also monitored to track the analytical reproducibility of isotope compositions. Isotope compositions for standard reference materials are shown in the table 1. 

```{r echo=FALSE}
# fit a line to the nitrogen data
N_fit = nitrogen %>% 
  filter(sample %in% standards_used) %>% 
  left_join(standards, by = 'sample') %>% 
  lm(d15N_true ~ d15N_measured , data = .) 

# fit a line to the carbon data
C_fit = carbon %>% 
  filter(sample %in% standards_used) %>% 
  left_join(standards, by = 'sample') %>% 
  lm(d13C_true ~ d13C_measured , data = .)

# display a table with the acceted values for the standards
knitr::kable(standards %>% filter(sample %in% standards_used),
             col.names = c('Standard', '$\\delta^{13}C$', '$\\delta^{15}N$', '$C \\%$', '$N \\%$'),
             caption = 'Accepted values for isotope standard reference materials')

```

```{r carbon_fit, echo=FALSE, fig.align='center', fig.height=4, fig.width=5}
# plot the results for carbon
carbon %>% 
  filter(sample %in% standards_used) %>% 
  left_join(standards, by = 'sample') %>% 
  with(plot(x = d13C_measured, 
            y = d13C_true, 
            pch = 21, 
            bg = 'tomato', 
            cex = 1.5,
            xlab = expression(delta^{13}~'C'['measured']),
            ylab = expression(delta^{13}~'C'['accepted']),
            main = expression(delta^{13}~'C standard calibration'),
            tcl = 0.25))
grid()
abline(v = 0)
abline(h = 0)
# add the calibration curve
abline(C_fit, lwd = 2)
```
A linear regression of the measured and accepted $\delta^{13}C$ values for the __`r paste(standards_used, collapse = ' and ')`__ standard reference materials returned a slope of __`r round(as.numeric(C_fit$coefficients[2]),3)`__ and an intercept of __`r round(as.numeric(C_fit$coefficients[1]),3)`__. with an $r^2$ of __`r summary(C_fit)$r.squared`__

```{r nitrogen_fit, echo=FALSE, fig.align='center', fig.height=4, fig.width=5}
# plot the results for nitrogen
nitrogen %>% 
  filter(sample %in% standards_used) %>% 
  left_join(standards, by = 'sample') %>% 
  with(plot(x = d15N_measured, 
            y = d15N_true, 
            pch = 21, 
            bg = 'tomato', 
            cex = 1.5,
            xlab = expression(delta^{15}~'N'['measured']),
            ylab = expression(delta^{15}~'N'['true']),
            main = expression(delta^{15}~'N standard calibration'),
            tcl = 0.25))
grid()
abline(v = 0)
abline(h = 0)
# add the calibration curve
abline(N_fit, lwd = 2)
```
A linear regression of the measured and accepted $\delta^{15}N$ values for the __`r paste(standards_used, collapse = ' and ')`__ standard reference materials returned a slope of __`r round(as.numeric(N_fit$coefficients[2]),3)`__ and an intercept of __`r round(as.numeric(N_fit$coefficients[1]),3)`__. with an $r^2$ of __`r summary(N_fit)$r.squared`__

```{r echo=FALSE}
### Correct the data using the linear regression results
N_fit %<>% coefficients()
C_fit %<>% coefficients()
nitrogen %<>% mutate(d15N_corrected = N_fit[1] + N_fit[2] * d15N_measured)
carbon %<>% mutate(d13C_corrected = C_fit[1] + C_fit[2] * d13C_measured)
```

# Standard Summary
Summary statistics for the __`r paste(standards_used, collapse = ' and ')`__ standard reference materials and the __`r paste(reference_material, collapse = ' and ')`__ in house reference are shown in the table 2. 
```{r echo=FALSE}
nitrogen %>% left_join(carbon, by = 'sample') %>% 
  filter(sample %in% c(standards_used, 
                       reference_material)) %>% 
  group_by(sample) %>% 
  summarise(d15N_mean = mean(d15N_corrected),
            d15N_sd = sd(d15N_corrected),
            d13C_mean = mean(d13C_corrected),
            d13C_sd = sd(d13C_corrected),
            N_wt_percent_mean = mean(N_wt_percent),
            N_wt_percent_sd = sd(N_wt_percent),
            C_wt_percent_mean = mean(C_wt_percent),
            C_wt_percent_sd = sd(C_wt_percent)) %>% 
  knitr::kable(digits = 2, 
               col.names = c('sample', 
                             '$\\delta^{15}N_{mean}$', 
                             '$\\delta^{15}N_{sd}$', 
                             '$\\delta^{13}C_{mean}$', 
                             '$\\delta^{13}C_{sd}$', 
                             '$N \\%_{mean}$',
                             '$N \\%_{sd}$',
                             '$C \\%_{mean}$',
                             '$C \\%_{sd}$'),
               caption = 'Summary statistics for standard 
              reference materials and in house reference materials.')
```

# Final Data Table
```{r echo = FALSE}
# render a table of the final data
final_data <- nitrogen %>% 
  left_join(carbon, 
            by = 'Identifier_1') %>% 
  select(Identifier_1,
         sample = sample.x,
         amount = amount.x,
         `Area 28`,
         `Area 44`,
         d13C_corrected,
         d15N_corrected,
         N_wt_percent,
         C_wt_percent) %>% 
  mutate(`C/N` = C_wt_percent/N_wt_percent)

final_data %>%
  select(Identifier_1, sample, d13C_corrected, d15N_corrected, N_wt_percent, C_wt_percent, `C/N`) %>%
  arrange(sample) %>% 
  knitr::kable(digits = 2, 
               col.names = c('Identifier 1',
                             'sample',
                             '$\\delta^{13}C$', 
                             '$\\delta^{15}N$', 
                             '$N \\%$',
                             '$C \\%$',
                             '$C:N$'),
               caption = 'Corrected isotope compositions, 
               total carbon and nitrogen content, and C:N ratios.')
final_data %>% arrange(sample) %>% write_csv(path = paste0(file_name,'_corrected.csv'))
```