---
set_title: 'EA Report'
author: "Robin B. Trayler"
date: "`r format(Sys.time(), '%B, %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{css, echo=FALSE}
.block-format {
  background-color: lightgrey2;
  border: 2px solid black;
  font-weight: bold;
}

/* Whole document: */
  body{
    font-family: Helvetica;
    font-size: 12pt;
  }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The $\delta^{13}C$ and $\delta^{15}N$ values and elemental carbon and nitrogen contents  of all samples were measured using a **Costech 4010** Elemental Analyzer coupled with a **Delta V Plus Continuous Flow Isotope Ratio Mass Spectrometer** housed in the Stable Isotope Facility at the University of California, Merced. 

This report includes several files:

1) A `.html` or `.pdf` report summarizing all data corrections and instrument reproducibility
2) A `.csv` file with the corrected data
3) A `.csv` file with the uncorrected data 
4) A `.Rmd` (R Markdown) document containing the **R** code used to correct the data

If you would like the original Isodat 3.0 `.dxf` files please request by sending an email to stableisotopes@ucmerced.edu


```{r, message = FALSE, echo = FALSE}
library(tidyverse)
library(viridis)
library(cowplot)
library(lubridate)
theme_set(theme_bw()) # set ggplot theme
```

```{r, echo = FALSE}
z_score <- function(x) {
  abs(x - mean(x)) / sd(x)
}
```

```{r, echo = FALSE, message = FALSE}
# standards <- read_csv(file = './data/EA_standards.csv')
standards <- tribble(~id2, ~d13C_true, ~d15N_true, ~C_percent, ~N_percent,
                     'USGS 40', -26.39, -4.52, 40.8, 9.52,
                     'USGS 41a', 36.55, 47.55, 40.8, 9.52,
                     'costech acetanilide', NA, NA, 71.09, 10.36) 

colors <- viridis(n = length(standards$id2) + 1, end = 0.8, option = 'D')
names(colors) <- c(standards$id2, TRUE)
```

# Setup {.tabset}
## File Path
```{r}
# file path to the isodat csv file 
file_path <- '~/Box Sync/Data Repository/EA/EA20210129.csv'
```

```{r, echo = FALSE}
# set the document file name to the name of the data file read in
file_name = tools::file_path_sans_ext(basename(file_path))
current_year <- format(Sys.Date(), '%Y')

#Check if directories exists and create if they doesn't
if(!dir.exists(path = paste0('~/Box Sync/Data Repository/EA/Corrected/',current_year))) {
  dir.create(path = paste0('~/Box Sync/Data Repository/EA/Corrected/', current_year))
}

if(!dir.exists(path = paste0('~/Box Sync/Data Repository/EA/Corrected/',current_year,'/', file_name))) {
  dir.create(path = paste0(path = paste0('~/Box Sync/Data Repository/EA/Corrected/',current_year,'/', file_name)))
}
```

## Analysis Parameters
### Select Sample Peaks
```{r}
# Which peak is the N2 analyte?
nitrogen_peak <- 3
# Which peak is the CO2 analyte?
carbon_peak <- 4
# How many total peaks should there be?
total_peaks <- 6
```

### Set Rejection Criteria
```{r}
# rejection criteria
small_peaks <- 500 # warn about peaks smaller with an amplitude smaller than this
# should 2 sigma outlier standards be dropped from isotope corrections
drop_outliers <- FALSE 
```

### Select Correction Standards
```{r}
# standards used in run
standards_used = c('USGS 41a', 'USGS 40')
# material to use for linearity corrections
linearity_material = c('USGS 40')
# material to use for drift corrections
drift_material = c('USGS 41a', 'USGS 40')
# material to use for weight calculations
weight_percent_material = c('USGS 40', 'USGS 41a', 'costech acetanilide')
# in house 'check' materials for run quality
check_material = c('EA sed', 'MB squid', 'peach', 'costech acetanilide', 'EA gelatin', 'Mb squid')
```

## Drift and Linearity Model Specification
### Instrument Drift
```{r}
# should a drift correction be applied to nitrogen
nitrogen_drift <- TRUE
# model choices are either 'lm' for linear or 'loess' for loess
nitrogen_drift_model <- 'lm'

# should a drift correction be applied to carbon
carbon_drift <- TRUE
# model choices are either 'lm' for linear or 'loess' for loess
carbon_drift_model <- 'lm'
```

### Mass Linearity
```{r}
# should a linearity correction be applied?
# mass linearity corrections use a linear model
nitrogen_linearity <- TRUE
carbon_linearity <- TRUE
```

```{r, echo = FALSE}
# set the document file name to the name of the data file read in
file_name = tools::file_path_sans_ext(basename(file_path))
```

---
title: "`r paste(file_name,' Report')`"
---

# Data Screening {.tabset}
```{r, message = FALSE, echo = FALSE}
# read data
raw_data <- read_csv(file = file_path)

data <- raw_data %>% 
  select(date = Date,
         time = Time,
         seq_nr = Row,
         id1 = `Identifier 1`, 
         id2 = `Identifier 2`,
         peak_nr = `Peak Nr.`,
         weight = Amount,
         area28 = `Area 28`,
         amp28 = `Ampl. 28`,
         amp44 = `Ampl. 44`,
         area44 = `Area 44`,
         d13C_measured = `d 13C/12C`,
         d15N_measured = `d 15N/14N`) %>% 
  mutate(date_time = mdy(date) + hms(time))
```

```{r, echo = FALSE}
# separate out the carbon data ------------------------------------------------
carbon <- data %>% 
  filter(peak_nr == carbon_peak) %>% 
  full_join(standards, by = 'id2') %>% 
  group_by(id2) %>% 
  mutate(d13C_dev = d13C_measured - mean(d13C_measured)) %>% 
  ungroup() %>% 
  mutate(weight_carbon = weight * C_percent / 100)
```

```{r, echo = FALSE}
# separate out the nitrogen data ----------------------------------------------
nitrogen <- data %>% 
  filter(peak_nr == nitrogen_peak) %>% 
  full_join(standards, by = 'id2') %>% 
  group_by(id2) %>% 
  mutate(d15N_dev = d15N_measured - mean(d15N_measured)) %>% 
  ungroup() %>% 
  mutate(weight_nitrogen = weight * N_percent / 100)
```

## Check for Missing Peaks
```{r, echo = FALSE, message = FALSE}
# check for missing peaks
missing <- data %>% 
  group_by(id1) %>% 
  summarize(n = !(n() == total_peaks))
```
Sample(s) with an Identifier 1 of __`r missing$id1[missing$n]`__ have fewer peaks than expected. Please check the raw data.

## Check for Small Peaks
```{r, echo = FALSE}
small_N <- data %>% 
  filter(peak_nr %in% nitrogen_peak & amp28 < small_peaks) %>% 
  select(id1)

small_C <- data %>% 
  filter(peak_nr %in% carbon_peak & amp44 < small_peaks) %>% 
  select(id1)
```

Sample(s) with an Identifier 1 of __`r small_N$id1`__ have a mass 28 (nitrogen) amplitude smaller than __`r small_peaks`__ mV. Data from these samples may be unreliable.

Sample(s) with an Identifier 1 of __`r small_C$id1`__ have a mass 44 (carbon) amplitude smaller than __`r small_peaks`__ mV. Data from these samples may be unreliable.

# Drift Correction {.tabset}
```{r, echo=FALSE}
# figure out axis limits ------------------------------------------------------
C_ylims <- carbon %>% 
  filter(id2 %in% drift_material) %>% 
  select(d13C_dev) %>% 
  range() %>% pretty() %>% range()

N_ylims <- nitrogen %>% 
  filter(id2 %in% drift_material) %>% 
  select(d15N_dev) %>% 
  range() %>% pretty() %>% range()
```

```{r, echo = FALSE, message=FALSE}
# carbon pre calibration plot -------------------------------------------------
C_drift_uncorrected <- carbon %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(mapping = aes(x = seq_nr,
                       y = d13C_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = carbon_drift_model,
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors) + 
  xlab('sequence number') + 
  ylab(expression(delta^13*C[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = C_ylims) + 
  guides(color=guide_legend(ncol=2))
```

```{r, echo = FALSE, message=FALSE}
# pre calibration plot --------------------------------------------------------
N_drift_uncorrected <- nitrogen %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(mapping = aes(x = seq_nr,
                       y = d15N_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = nitrogen_drift_model,
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors) + 
  xlab('sequence number') + 
  ylab(expression(delta^15*N[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = N_ylims) + 
  guides(color=guide_legend(ncol=2))
```


```{r, message = FALSE, echo = FALSE}
# fit a Drift Correction ------------------------------------------------------
if(carbon_drift) {
  if(carbon_drift_model == 'lm') {
    C_drift_fit <- carbon %>% 
      filter(id2 %in% drift_material) %>%
      lm(d13C_dev ~ seq_nr, data = .)
  } else if(carbon_drift_model == 'loess') {
    C_drift_fit <- carbon %>% 
      filter(id2 %in% drift_material) %>%
      loess(d13C_dev ~ seq_nr, data = .)
  }
  carbon <- carbon %>% 
    mutate(d13C_measured = d13C_measured - predict(C_drift_fit, newdata = carbon)) %>%
    group_by(id2) %>%
    mutate(d13C_dev = d13C_measured - mean(d13C_measured)) %>%
    ungroup()
}
```

```{r, message = FALSE, echo = FALSE}
# fit a Drift Correction ------------------------------------------------------
if(nitrogen_drift) {
  if(nitrogen_drift_model == 'lm') {
    N_drift_fit <- nitrogen %>% 
      filter(id2 %in% drift_material) %>%
      lm(d15N_dev ~ seq_nr, data = .)
  } else if(nitrogen_drift_model == 'loess') {
    N_drift_fit <- nitrogen %>% 
      filter(id2 %in% drift_material) %>%
      loess(d15N_dev ~ seq_nr, data = .)
  }
  
  nitrogen <- nitrogen %>% 
    mutate(d15N_measured = d15N_measured - predict(N_drift_fit, newdata = nitrogen)) %>%
    group_by(id2) %>%
    mutate(d15N_dev = d15N_measured - mean(d15N_measured)) %>%
    ungroup()
}
```

```{r, echo = FALSE, message=FALSE}
if(carbon_drift){
  # post correction plot --------------------------------------------------------
  C_drift_corrected <- carbon %>% 
    filter(id2 %in% drift_material) %>% 
    ggplot(mapping = aes(x = seq_nr,
                         y = d13C_dev,
                         color = id2)) + 
    geom_point(size = 3) + 
    theme(legend.position = 'top') + 
    geom_smooth(method = 'lm',
                se = FALSE, 
                color = 'black') + 
    scale_color_manual(values = colors) + 
    xlab('sequence number') + 
    ylab(expression(delta^13*C[deviation])) + 
    ggtitle('Corrected') + 
    coord_cartesian(ylim = C_ylims) + 
    guides(color=guide_legend(ncol=2))
}

if(nitrogen_drift){
  # post correction plot --------------------------------------------------------
  N_drift_corrected <- nitrogen %>% 
    filter(id2 %in% drift_material) %>% 
    ggplot(mapping = aes(x = seq_nr,
                         y = d15N_dev,
                         color = id2)) + 
    geom_point(size = 3) + 
    theme(legend.position = 'top') + 
    geom_smooth(method = 'lm',
                se = FALSE, 
                color = 'black') + 
    scale_color_manual(values = colors) + 
    xlab('sequence number') + 
    ylab(expression(delta^15*N[deviation])) + 
    ggtitle('Corrected') + 
    coord_cartesian(ylim = N_ylims) + 
    guides(color=guide_legend(ncol=2))
}
```

## Nitrogen Drift Correction
Nitrogen isotope compositional drift over the period of analysis was monitored using the **`r paste(drift_material, collapse = ' and ')`** reference material(s). Since a drift correction was `r ifelse(carbon_drift,' ','not')` requested, A **`r if(nitrogen_drift_model == 'lm') {paste('linear')} else if(nitrogen_drift_model == 'loess') {paste('loess')}`** model was `r ifelse(nitrogen_drift,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
if(nitrogen_drift){
  plot_grid(N_drift_uncorrected, N_drift_corrected)
}
```


## Carbon Drift Correction
Carbon isotope compositional drift over the period of analysis was monitored using the **`r paste(drift_material, collapse = ' and ')`** reference material(s). Since a drift correction was `r ifelse(carbon_drift,' ','not')` requested, A **`r if(carbon_drift_model == 'lm') {paste('linear')} else if(carbon_drift_model == 'loess') {paste('loess')}`** model was `r ifelse(carbon_drift,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
if(carbon_drift){
  plot_grid(C_drift_uncorrected, C_drift_corrected)
}
```

# Linearity Corrections {.tabset}
```{r, echo = FALSE}
N_ylims <- nitrogen %>% 
  filter(id2 %in% linearity_material) %>% 
  select(d15N_dev) %>% 
  range() %>% pretty() %>% range()

C_ylims <- carbon %>% 
  filter(id2 %in% linearity_material) %>% 
  select(d13C_dev) %>% 
  range() %>% pretty() %>% range()
```

```{r, echo = FALSE}
# Carbon
# pre calibration plot --------------------------------------------------------
C_lin_uncorrected <- carbon %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area44,
                       y = d13C_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors) + 
  xlab('mass 44 peak area (mV/s)') + 
  ylab(expression(delta^13*C[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = C_ylims) + 
  guides(color=guide_legend(ncol=2))

# Nitrogen
# pre calibration plot --------------------------------------------------------
N_lin_uncorrected <- nitrogen %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area28,
                       y = d15N_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors) + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab(expression(delta^15*N[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = N_ylims) + 
  guides(color=guide_legend(ncol=2))
```

```{r, echo = FALSE, message = FALSE}
# Carbon
# fit a linearity correction --------------------------------------------------
C_lin_fit <- carbon %>% 
  filter(id2 %in% linearity_material) %>% 
  lm(formula = d13C_dev ~ area44, data = .)
C_lin_sum <- summary(C_lin_fit)

if(carbon_linearity) {
  carbon <- carbon %>% 
    mutate(d13C_measured = d13C_measured - predict(C_lin_fit, newdata = carbon)) %>% 
    group_by(id2) %>% 
    mutate(d13C_dev = d13C_measured - mean(d13C_measured)) %>% 
    ungroup()
}

# Nitrogen
# fit a linearity correction --------------------------------------------------
N_lin_fit <- nitrogen %>% 
  filter(id2 %in% linearity_material) %>% 
  lm(formula = d15N_dev ~ area28, data = .)
N_lin_sum <- summary(N_lin_fit)

if(nitrogen_linearity){
  nitrogen <- nitrogen %>% 
    mutate(d15N_measured = d15N_measured - predict(N_lin_fit, newdata = nitrogen)) %>% 
    group_by(id2) %>% 
    mutate(d15N_dev = d15N_measured - mean(d15N_measured)) %>% 
    ungroup()
}
```

```{r, echo = FALSE}
# post calibration plot -------------------------------------------------------
C_lin_corrected <- carbon %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area44,
                       y = d13C_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors) + 
  xlab('mass 44 peak area (mV/s)') + 
  ylab(expression(delta^13*C[deviation])) + 
  ggtitle('Corrected') + 
  coord_cartesian(ylim = C_ylims) + 
  guides(color=guide_legend(ncol=2))

N_lin_corrected <- nitrogen %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area28,
                       y = d15N_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors) + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab(expression(delta^15*N[deviation])) + 
  ggtitle('Corrected') + 
  coord_cartesian(ylim = N_ylims) + 
  guides(color=guide_legend(ncol=2))
```

## Nitrogen Linearity Correction
Mass linearity of nitrogen isotope compositions  was monitored using the **`r paste(linearity_material, collapse = ' and ')`** reference material(s). Since a linearity correction was `r ifelse(nitrogen_linearity,' ','not')` requested, A linear trend with an $r^2$ of **`r round(N_lin_sum$r.squared, 2)`** was `r ifelse(nitrogen_linearity,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center'}
if(nitrogen_linearity) {
  plot_grid(N_lin_uncorrected, N_lin_corrected)
}
```

## Carbon Linearity Correction
Mass linearity of carbon isotope compositions  was monitored using the **`r paste(linearity_material, collapse = ' and ')`** reference material(s). Since a linearity correction was `r ifelse(carbon_linearity,' ','not')` requested, A linear trend with an $r^2$ of **`r round(C_lin_sum$r.squared, 2)`** was `r ifelse(carbon_linearity,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center'}
if(carbon_linearity) {
  plot_grid(C_lin_uncorrected, C_lin_corrected)
}
```

# Standardization {.tabset}
```{r, echo = FALSE, message=FALSE, fig.align='center'}
# calibration plot
C_scale <- carbon %>% 
  filter(id2 %in% standards_used) %>% 
  ggplot(mapping = aes(x = d13C_measured,
                       y = d13C_true,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors) +
  geom_smooth(method = 'lm', aes(group = 1), color = 'black') + 
  theme(legend.position = 'top') + 
  xlab(expression(delta^13*C[measured])) + 
  ylab(expression(delta^13*C[accepted])) + 
  guides(color=guide_legend(ncol=2))

N_scale <- nitrogen %>% 
  filter(id2 %in% standards_used) %>% 
  ggplot(mapping = aes(x = d15N_measured,
                       y = d15N_true,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors) +
  geom_smooth(method = 'lm', aes(group = 1), color = 'black') + 
  theme(legend.position = 'top') + 
  xlab(expression(delta^15*N[measured])) + 
  ylab(expression(delta^15*N[accepted])) + 
  guides(color=guide_legend(ncol=2))

plot_grid(N_scale, C_scale)
```

```{r, echo = FALSE}
if(drop_outliers){
  C_std_fit <- carbon %>%  
    filter(id2 %in% standards_used) %>% 
    group_by(id2) %>% 
    mutate(z_score = z_score(d13C_measured)) %>% 
    filter(z_score < 2) %>% 
    ungroup() %>% 
    lm(formula = d13C_true ~ d13C_measured, data = .)
  
  N_std_fit <- nitrogen %>%  
    filter(id2 %in% standards_used) %>% 
    group_by(id2) %>% 
    mutate(z_score = z_score(d15N_measured)) %>% 
    filter(z_score < 2) %>% 
    ungroup() %>% 
    lm(formula = d15N_true ~ d15N_measured, data = .)
  carbon <- carbon %>% 
    mutate(d13C_corrected = predict(C_std_fit, newdata = carbon))
  nitrogen <- nitrogen %>% 
    mutate(d15N_corrected = predict(N_std_fit, newdata = nitrogen))
  
} else {
  C_std_fit <- carbon %>%  
    filter(id2 %in% standards_used) %>% 
    lm(formula = d13C_true ~ d13C_measured, data = .)
  N_std_fit <- nitrogen %>%  
    filter(id2 %in% standards_used) %>% 
    lm(formula = d15N_true ~ d15N_measured, data = .) 
  carbon <- carbon %>% 
    mutate(d13C_corrected = predict(C_std_fit, newdata = carbon))
  nitrogen <- nitrogen %>% 
    mutate(d15N_corrected = predict(N_std_fit, newdata = nitrogen))
}

```

```{r, echo = FALSE}
N_r2 <- N_std_fit %>% summary() %>% pluck('r.squared')

C_r2 <- C_std_fit %>% summary() %>% pluck('r.squared')
```

The isotope composition of unknowns samples were corrected to the international PDB ($\delta^{13}C$) and Air ($\delta^{15}N$) scales using the **`r paste(standards_used, collapse = ' and ')`** standard reference materials. Isotope compositions for standard reference materials are shown in the **Table 1**.

```{r, echo = FALSE}
standards %>% 
  filter(id2 %in% standards_used) %>%
  select(id2, 
         d13C_true, 
         d15N_true) %>% 
  knitr::kable(col.names = c('standard', 
                             '$\\delta^{13}C$', 
                             '$\\delta^{15}N$'),
               caption = 'Table 1: Certified isotope compositions for standard reference materials')
```

# Elemental Content {.tabset}
```{r echo=FALSE, fig.align='center', message=FALSE}
C_weight <- carbon %>% 
  filter(id2 %in% weight_percent_material) %>% 
  ggplot(mapping = aes(x = area44, 
                       y = weight_carbon,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors) + 
  theme(legend.position = 'top') + 
  geom_smooth(aes(group = 1), color = 'black', method = 'lm') + 
  xlab('mass 44 peak area (mV/s)') + 
  ylab('carbon content (mg)') + 
  guides(color=guide_legend(ncol=2))

N_weight <- nitrogen %>% 
  filter(id2 %in% weight_percent_material) %>% 
  ggplot(mapping = aes(x = area28, 
                       y = weight_nitrogen,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors) + 
  theme(legend.position = 'top') + 
  geom_smooth(aes(group = 1), color = 'black', method = 'lm') + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab('nitrogen content (mg)') + 
  guides(color=guide_legend(ncol=2))

plot_grid(N_weight, C_weight)
```

```{r, echo = FALSE}
C_wt_fit <- carbon %>% 
  filter(id2 %in% weight_percent_material) %>% 
  lm(formula = weight_carbon ~ area44, data = .)
carbon <- carbon %>% 
  mutate(C_wt_percent = predict(C_wt_fit, newdata = carbon) / weight * 100)

N_wt_fit <- nitrogen %>% 
  filter(id2 %in% weight_percent_material) %>% 
  lm(formula = weight_nitrogen ~ area28, data = .)
nitrogen <- nitrogen %>% 
  mutate(N_wt_percent = predict(N_wt_fit, newdata = nitrogen) / weight * 100)
```
Total carbon and nitrogen contents were calculated using the **`r paste(weight_percent_material, collapse = ' and ')`** reference materials. Certified carbon and nitrogen contents for these reference materials are shown in the **Table 2**. 

```{r, echo = FALSE}
standards %>% 
  filter(id2 %in% weight_percent_material) %>%
  select(id2, 
         C_percent, 
         N_percent) %>% 
  knitr::kable(col.names = c('standard', 
                             '$C \\%$', 
                             '$N \\%$'),
               caption = 'Table 2: Certified elemental contents for standard reference materials')
```

# Final Data {.tabset}
## Summary 

```{r, echo = FALSE, message = FALSE}
carbon <- carbon %>% 
  select(
    id1, 
    id2, 
    weight,
    area44, 
    d13C_corrected,
    C_wt_percent)

nitrogen <- nitrogen %>% 
  select(
    id1, 
    id2, 
    weight,
    area28, 
    d15N_corrected,
    N_wt_percent)

final_data <- full_join(carbon, nitrogen) %>% 
  select(Identifier_1 = id1, 
         sample = id2, 
         amount = weight,
         `Area 28` = area28,
         `Area 44`= area44,
         d13C_corrected,
         d15N_corrected,
         N_wt_percent,
         C_wt_percent) %>% 
  mutate(`C/N` = C_wt_percent/N_wt_percent)
```

```{r, echo = FALSE, message = FALSE}
final_data %>% 
  filter(sample %in% c(standards_used, check_material, drift_material, linearity_material)) %>% 
  group_by(sample) %>% 
  summarise(d15N_mean = mean(d15N_corrected),
            d15N_sd = sd(d15N_corrected),
            d13C_mean = mean(d13C_corrected),
            d13C_sd = sd(d13C_corrected),
            N_wt_percent_mean = mean(N_wt_percent),
            N_wt_percent_sd = sd(N_wt_percent),
            C_wt_percent_mean = mean(C_wt_percent),
            C_wt_percent_sd = sd(C_wt_percent),
            n = n()) %>% 
  knitr::kable(digits = 2, 
               col.names = c('sample', 
                             '$\\delta^{15}N_{mean}$', 
                             '$\\delta^{15}N_{sd}$', 
                             '$\\delta^{13}C_{mean}$', 
                             '$\\delta^{13}C_{sd}$', 
                             '$N \\%_{mean}$',
                             '$N \\%_{sd}$',
                             '$C \\%_{mean}$',
                             '$C \\%_{sd}$',
                             'n'),
               caption = 'Summary statistics for standard 
              reference materials and in house reference materials.')
```

## Full Data Table
```{r, echo = FALSE}
final_data %>%
  arrange(sample) %>%
  mutate_if(is.numeric, round, 3) %>% 
  knitr::kable(col.names = c('Identifier 1', 
                             'sample',
                             'amount (mg)', 
                             'area 28',
                             'area 44',
                             '$\\delta^{13}C$',
                             '$\\delta^{15}N$',
                             '$N \\%$',
                             '$C \\%$', 
                             '$C \\ /N$'),
               caption = 'Corrected isotope compositions, 
               total carbon and nitrogen content, and C/N ratios. C/N ratios are reported as wt/wt.')
```

```{r, echo = FALSE, message = FALSE}
final_data %>% 
  arrange(sample) %>% 
  write_csv(path = paste0('~/Box Sync/Data Repository/EA/Corrected/',current_year,'/',file_name,'/',file_name,'_corrected.csv'))

raw_data %>% 
  write_csv(path = paste0('~/Box Sync/Data Repository/EA/Corrected/',current_year,'/',file_name,'/',file_name, '_raw.csv'))

file.copy(from = 'EA_tidy_reduction.Rmd', to = paste0('~/Box Sync/Data Repository/EA/Corrected/',current_year,'/',file_name,'/',file_name,'_code.Rmd'))
```
