---
title: 'GB20190820 Report'
author: "Robin B. Trayler"
date: "8/21/2019"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(magrittr)
```

The $\delta^{13}C$ and $\delta^{18}O$ values of all samples were measured using a __Thermo-Finnigan Gasbench II__ coupled with a __Delta V Plus Continuous Flow Isotope Ratio Mass Spectrometer__ housed in the Stable Isotope Facility at the University of California, Merced. If you would like the original Isodat 3.0 output file please request by sending an email to StableIsotopes@ucmerced.edu

# Run Parameters
```{r define_parameters, echo = TRUE}
####################################################################
file_path = "~/Desktop/Gasbench/GB20190820.csv" # location of raw data file #
####################################################################
file_name = tools::file_path_sans_ext(basename(file_path))
reference_peaks = 3 # first three peaks are references
sample_peaks = 4 # samples start at peak 4
total_peaks = 10 # total number of peaks is 10
standards_used = c("'USGS 44'", "'CM'", "'NBS 18'") # standards used in run
reference_material = "'NIST 120C'" # reference material
linearity_correction = FALSE # should linearity correction be performed?
```

```{r, echo = FALSE}
# define values for standards
standards <- 
  tibble(sample = c('NBS 18', 'NBS 19', 'USGS 44', 'CM'), 
         d13C_true = c(-5.014, 1.95, -42.1, 2.10),
         d18O_true = c(-23.2, -2.20, NA, -2.01))
```

```{r load_data, echo=FALSE, message=FALSE, warning=FALSE}
# Load data
data = read_csv(file = file_path)
#filter out the useful columns
data %<>% select(Identifier_1 = `Identifier 1`,
                 sample = `Identifier 2`,
                 amount = Comment,
                 `Area 44` = `Area 44`,
                 d13C_measured = `d 13C/12C`,
                 d18O_measured = `d 18O/16O`,
                 `Peak Nr` = `Peak Nr.`)
```

# Check For Bad Samples
```{r check_data, echo = FALSE}
# Check for missing peaks
ID = unique(data[['Identifier_1']])

bad_peaks = rep(FALSE, length = length(ID))
for (i in 1:length(ID)) {
  if (data %>% filter(Identifier_1 == ID[i]) %>% select(`Peak Nr`) %>% nrow() < total_peaks) {
    bad_peaks[i] = TRUE
  }
}
```

Sample(s) with an Identifier 1 of __`r ID[bad_peaks]`__ have fewer peaks than expected. Please check the raw data. 

```{r select_standards,echo=FALSE}
# build a string of standard values for use in filtering the data later
standard_string = character()
for (i in 1:length(standards_used)) {
  if (i >= 1 & i < length(standards_used)) {
    standard_string = paste(standard_string, 'sample == ', standards_used[i], '|')
  }
  else
    if (i == length(standards_used)) {
      standard_string = paste(standard_string, 'sample == ', standards_used[i])
    }
}
# add the reference material to the string
reference_string = paste(standard_string,'| sample ==', reference_material)
```

```{r, echo=FALSE}
# filter and select
#Filter out the nitrogen and carbon data seperately. This is based on peak number which can cause problems if there are  #missing peaks. It might be better to filter on retention time so the code will only grab peaks within the correct time #window.
sample_data = data %>% 
  filter(`Peak Nr` >= sample_peaks) %>% 
  select(Identifier_1, 
         sample,
         amount,
         `Area 44`,
         d13C_measured,
         d18O_measured,
         `Peak Nr`)
```

```{r linearity, echo=FALSE}
# Linearity corrections
#Still working on this. Right now its relies on sample names starting with 'lin' to indicate that they should be used to do #a linearity correction.
```

```{r echo = FALSE}
sample_data %<>% 
  group_by(Identifier_1) %>% 
  summarize(sample = unique(sample),
            amount = unique(amount),
            d13C = mean(d13C_measured),
            d13C_sd = sd(d13C_measured),
            d18O = mean(d18O_measured),
            d18O_measured = sd(d18O_measured))
```
# Standard correction
The isotope composition of unknowns samples were corrected to the international PDB scale using the __`r paste(standards_used, collapse = ' and ')`__ standard reference materials. An in house reference material, __`r paste(reference_material)`__ was also monitored to track the analytical reproducibility of isotope compositions. Isotope compositions for standard reference materials are shown in the table 1. 

```{r echo=FALSE}
# fit a line to the nitrogen data
C_fit = sample_data %>% 
  filter(eval(parse(text = standard_string))) %>% 
  left_join(standards, by = 'sample') %>% 
  lm(d13C_true ~ d13C, data = .) 

O_fit = sample_data %>% 
  filter(eval(parse(text = standard_string))) %>% 
  left_join(standards, by = 'sample') %>% 
  lm(d18O_true ~ d18O, data = .) 

# # display a table with the acceted values for the standards
knitr::kable(standards %>% filter(eval(parse(text = standard_string))),
             col.names = c('Standard', '$\\delta^{13}C$', '$\\delta^{18}O$'),
             caption = 'Accepted values for isotope standard reference materials')
```

```{r carbon_fit, echo=FALSE, fig.align='center', fig.height=4, fig.width=5}
# plot the results for carbon
sample_data %>% 
  filter(eval(parse(text = standard_string))) %>% 
  left_join(standards, by = 'sample') %>% 
  with(plot(x = d13C, 
            y = d13C_true, 
            pch = 21, 
            bg = 'tomato', 
            cex = 1.5,
            xlab = expression(delta^{13}~'C'['measured']),
            ylab = expression(delta^{13}~'C'['accepted']),
            main = expression(delta^{13}~'C standard calibration'),
            tcl = 0.25))
grid()
abline(v = 0)
abline(h = 0)
# add the calibration curve
abline(C_fit, lwd = 2)
```
A linear regression of the measured and accepted $\delta^{13}C$ values for the __`r paste(standards_used, collapse = ' and ')`__ standard reference materials returned a slope of __`r round(as.numeric(C_fit$coefficients[2]),3)`__ and an intercept of __`r round(as.numeric(C_fit$coefficients[1]),3)`__. with an $r^2$ of __`r summary(C_fit)$r.squared`__

```{r nitrogen_fit, echo=FALSE, fig.align='center', fig.height=4, fig.width=5}
# plot the results for nitrogen
sample_data %>% 
  filter(eval(parse(text = standard_string))) %>% 
  left_join(standards, by = 'sample') %>% 
  with(plot(x = d18O, 
            y = d18O_true, 
            pch = 21, 
            bg = 'tomato', 
            cex = 1.5,
            xlab = expression(delta^{18}~'O'['measured']),
            ylab = expression(delta^{18}~'O'['accepted']),
            main = expression(delta^{18}~'O standard calibration'),
            tcl = 0.25))
grid()
abline(v = 0)
abline(h = 0)
# add the calibration curve
abline(O_fit, lwd = 2)
```
A linear regression of the measured and accepted $\delta^{18}O$ values for the __`r paste(standards_used, collapse = ' and ')`__ standard reference materials returned a slope of __`r round(as.numeric(O_fit$coefficients[2]),3)`__ and an intercept of __`r round(as.numeric(O_fit$coefficients[1]),3)`__. with an $r^2$ of __`r summary(O_fit)$r.squared`__

```{r echo=FALSE}
## Correct the data using the linear regression results
C_fit %<>% coefficients()
O_fit %<>% coefficients()
sample_data %<>% mutate(d13C_corrected = C_fit[1] + C_fit[2] * d13C)
sample_data %<>% mutate(d18O_corrected = O_fit[1] + O_fit[2] * d18O)
```

# Standard Summary
Summary statistics for the __`r paste(standards_used, collapse = ' and ')`__ standard reference materials and the __`r paste(reference_material, collapse = ' and ')`__ in house reference are shown in the table 2. 
```{r echo=FALSE}
sample_data %>%
  filter(eval(parse(text = reference_string))) %>% 
  group_by(sample) %>% 
  
  summarise(d13C_mean = mean(d13C_corrected),
            d13C_sd = sd(d13C_corrected),
            d18O_mean = mean(d18O_corrected),
            d18O_sd = sd(d18O_corrected)) %>% 
  knitr::kable(digits = 2, 
               col.names = c('sample', 
                             '$\\delta^{13}C_{mean}$', 
                             '$\\delta^{13}C_{sd}$', 
                             '$\\delta^{18}O_{mean}$', 
                             '$\\delta^{18}O_{sd}$'),
               caption = 'Summary statistics for standard 
              reference materials and in house reference materials.')
```

# Final Data Table
```{r echo = FALSE}
# render a table of the final data
final_data = sample_data %>% 
  select(Identifier_1, sample, amount, d13C_corrected, d18O_corrected) %>% 
  arrange(sample) 
final_data %>% 
  knitr::kable(digits = 2,
               col.names = c('Identifier 1', 
                             'sample', 'Amount (mg)',
                             '$\\delta^{13}C$', 
                             '$\\delta^{18}O$'),
               caption = 'Corrected isotope composition')
final_data %>% write_csv(path = paste0(file_name,'_corrected.csv'))
```
