---
set_title: 'GB Report'
author: "Robin B. Trayler"
date: "`r format(Sys.time(), '%B, %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The $\delta^{13}C$ and $\delta^{18}O$ values of all samples were measured using a **Thermo-Finnigan Gasbench II** coupled with a **Delta V Plus Continuous Flow Isotope Ratio Mass Spectrometer** housed in the Stable Isotope Facility at the University of California, Merced.

This report includes several files:

1)  A `.html` or `.pdf` report summarizing all data corrections and instrument reproducibility
2)  A `.csv` file with the corrected data
3)  A `.csv` file with the uncorrected data
4)  A `.Rmd` (R Markdown) document containing the **R** code used to correct the data

If you would like the original Isodat 3.0 `.dxf` files please request by sending an email to [stableisotopes\@ucmerced.edu](mailto:stableisotopes@ucmerced.edu){.email}

```{r, message = FALSE, echo = FALSE}
library(tidyverse)
library(viridis)
library(cowplot)
theme_set(theme_bw()) # set ggplot theme
```

```{r, echo = FALSE}
# define a z score function
z_score <- function(x) {
  abs(x - mean(x)) / sd(x)
}
```

```{r, echo = FALSE, message = FALSE}
# define standard values
standards <- tribble(~id2,     ~d13C_true, ~d18O_true,
                     'NBS 18',  -5.014,      -23.2,
                     'NBS 19',  1.95,        -2.2,
                     'USGS 44', -42.1,        NA,
                     'CM',      2.1,         -2.01,
                     'alpha low',-50.0,       -38,
                     'alpha mid', -39.7,      NA,
                     'alpha high', -3.3,     -5.6) 
colors <- viridis(n = length(standards$id2) + 1, end = 0.8)
names(colors) <- c(standards$id2, TRUE)
```

# Setup {.tabset}

## Set File Path

```{r}
# file path to the isodat csv file 
file_path <- '~/Box Sync/Data Repository/Gasbench/GB20220613.csv'
```

```{r, echo = FALSE}
# set the document file name to the name of the data file read in
file_name = tools::file_path_sans_ext(basename(file_path))
```

---
title: "`r paste(file_name,' Report')`"
---

## Analysis Parameters

### Select Sample Peaks

```{r}
# analyte peaks
sample_peaks <- 5:20
reference_peaks <- 1:3
```

### Set Rejection Criteria

```{r}
# rejection criteria
small_peaks <- 500 # warn about peaks smaller than ###mV

# should 2 sigma outlier standards be dropped?
drop_outliers <- FALSE
```

### Select Correction Standards

```{r}
# standards used in run
standards_used = c('NBS 18', 'CM', 'USGS 44')
# standards_used = c('alpha low', 'alpha high')
# material to use for linearity corrections
linearity_material = c('NBS 18', 'CM')
# linearity_material = c('alpha low', 'alpha high')
# material to use for drift corrections
drift_material = c('NBS 18', 'CM')
# drift_material = c('alpha low', 'alpha high')
# material to use for weight calculations
# weight_percent_material = c('NBS 18', 'USGS 44', 'CM')
# in house 'check' materials for run quality
check_material = c('NIST 120C')
# check_material = c('alpha mid')
```

## Drift and Linearity Model Specification

### Instrument Drift

```{r}
# should a drift correction be applied to carbon?
carbon_drift <- TRUE
# model choices are either 'lm' for linear or 'loess' for loess
carbon_drift_model <- 'lm'

# should a drift correction be applied to oxygen
oxygen_drift <- TRUE
# model choices are either 'lm' for linear or 'loess' for loess
oxygen_drift_model <- 'lm'
```

### Mass Linearity

```{r}
# should a linearity correction be applied?
# mass linearity corrections use a linear model
carbon_linearity <- TRUE
oxygen_linearity <- TRUE
```

# Data Screening

```{r, message = FALSE, echo = FALSE}
# read data
raw_data <- read_csv(file = file_path)
data <- raw_data %>% 
  select(
    seq_nr = Row,
    id1 = `Identifier 1`, 
    id2 = `Identifier 2`,
    peak_nr = `Peak Nr.`,
    amp44 = `Ampl. 44`,
    area44 = `Area 44`,
    d13C_measured = `d 13C/12C`,
    d18O_measured = `d 18O/16O`) %>%
  filter(peak_nr %in% sample_peaks) %>% 
  group_by(id1) %>% 
  mutate(d13C_dev = d13C_measured - mean(d13C_measured),
         d18O_dev = d18O_measured - mean(d18O_measured)) %>% 
  ungroup() 
```

# Linearity Correction {.tabset}

```{r, echo = FALSE, message = FALSE}
# figure out limits for plots 
C_ylims <- data %>% 
  select(d13C_dev) %>% 
  range() %>% pretty() %>% range()

O_ylims <- data %>% 
  select(d18O_dev) %>% 
  range() %>% pretty() %>% range()
```

```{r, echo = FALSE, message = FALSE}
C_lin_uncorrected <- data %>%
  ggplot(aes(x = area44, y = d13C_dev, color = id2)) + 
  geom_point(size = 3, alpha = 0.75) + 
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = 'none') + 
  geom_smooth(aes(group = 1), 
              method = 'loess', 
              se = FALSE, 
              color = 'black') + 
  guides(color=guide_legend(ncol=2)) + 
  xlab('mass 44 peak area (mV/s)') + 
  ylab(expression(delta^13*C[deviation])) + 
  coord_cartesian(ylim = C_ylims) + 
  ggtitle('Uncorrected')

O_lin_uncorrected <- data %>%
  ggplot(aes(x = area44, y = d18O_dev, color = id2)) + 
  geom_point(size = 3, alpha = 0.75) + 
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = 'none') + 
  geom_smooth(aes(group = 1), 
              method = 'loess', 
              se = FALSE, 
              color = 'black') + 
  guides(color=guide_legend(ncol=2)) + 
  xlab('mass 44 peak area (mV/s)') + 
  ylab(expression(delta^18*O[deviation])) + 
  coord_cartesian(ylim = O_ylims) + 
  ggtitle('Uncorrected')
```

```{r, echo = FALSE, message = FALSE}
# apply the linearity correction
if(carbon_linearity) {
  C_lin_fit <- data %>% 
    loess(d13C_dev ~ area44, data = .)
  data <- data %>%
    mutate(d13C_measured = d13C_measured - predict(C_lin_fit, newdata = data)) %>% 
    group_by(seq_nr) %>% 
    mutate(d13C_dev = d13C_measured - mean(d13C_measured)) %>% 
    ungroup()
}
```

```{r, echo = FALSE, message = FALSE}
# apply the linearity correction
if(oxygen_linearity) {
  O_lin_fit <- data %>% 
    loess(d18O_dev ~ area44, data = .)
  
  data <- data %>%
    mutate(d18O_measured = d18O_measured - predict(O_lin_fit, newdata = data)) %>% 
    group_by(seq_nr) %>% 
    mutate(d18O_dev = d18O_measured - mean(d18O_measured)) %>% 
    ungroup()
}
```

```{r, echo = FALSE, message = FALSE}
C_lin_corrected <- data %>%
  ggplot(aes(x = area44, y = d13C_dev, color = id2)) + 
  geom_point(size = 3, alpha = 0.75) + 
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = 'none') + 
  geom_smooth(aes(group = 1), 
              method = 'loess', 
              se = FALSE, 
              color = 'black') + 
  guides(color=guide_legend(ncol=2)) + 
  xlab('mass 44 peak area (mV/s)') + 
  ylab(expression(delta^13*C[deviation])) + 
  coord_cartesian(ylim = C_ylims) + 
  ggtitle('Corrected')

O_lin_corrected <- data %>%
  ggplot(aes(x = area44, y = d18O_dev, color = id2)) + 
  geom_point(size = 3, alpha = 0.75) + 
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = 'none') + 
  geom_smooth(aes(group = 1), 
              method = 'loess', 
              se = FALSE, 
              color = 'black') + 
  guides(color=guide_legend(ncol=2)) + 
  xlab('mass 44 peak area (mV/s)') + 
  ylab(expression(delta^18*O[deviation])) + 
  coord_cartesian(ylim = O_ylims) + 
  ggtitle('Corrected')

```

## Carbon Linearity Correction

A locally smoothed polynomial (loess) was fit to mass 44 peak area (mV/s) versus the pooled within sample $\delta^{13}C$ deviations for all analyses. Since a linearity correction was `r ifelse(carbon_linearity,' ','not')` requested, the trend was `r ifelse(carbon_linearity,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
plot_grid(C_lin_uncorrected, C_lin_corrected)
```

## Oxygen Linearity Correction

A locally smoothed polynomial (loess) was fit to mass 44 peak area (mV/s) versus the pooled within sample $\delta^{18}O$ deviations for all analyses. Since a linearity correction was `r ifelse(oxygen_linearity,' ','not')` requested, the trend was `r ifelse(oxygen_linearity,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
plot_grid(O_lin_uncorrected, O_lin_corrected)
```

```{r, echo = FALSE, message = FALSE}
# aggregate the data 
data <-  data %>% 
  group_by(id1) %>% 
  summarize(id2 = unique(id2), 
            d13C_sd = sd(d13C_measured),
            d18O_sd = sd(d18O_measured),
            d13C_measured = mean(d13C_measured),
            d18O_measured = mean(d18O_measured),
            seq_nr = unique(seq_nr)) %>% 
  full_join(standards %>% filter(id2 %in% standards_used), by = 'id2') %>% 
  ungroup() %>% 
  group_by(id2) %>% 
  mutate(d13C_dev = d13C_measured - mean(d13C_measured),
         d18O_dev = d18O_measured - mean(d18O_measured)) %>% 
  ungroup()
```

# Drift Correction {.tabset}

```{r, echo = FALSE, message = FALSE}
# figure out limits for plots 
C_ylims <- data %>% 
  select(d13C_dev) %>% 
  range() %>% pretty() %>% range()

O_ylims <- data %>% 
  select(d18O_dev) %>% 
  range() %>% pretty() %>% range()
```

```{r, echo = FALSE, message=FALSE}
# Carbon drift plot
C_drift_uncorrected <- data %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(aes(x = seq_nr, 
             y = d13C_dev,
             color = id2)) + 
  geom_point(size = 3) + 
  geom_smooth(method = 'lm', 
              formula = y ~ x, 
              se = FALSE,
              mapping = aes(group = 1),
              color = 'black') + 
  scale_color_manual(values = colors) + 
  coord_cartesian(ylim = C_ylims) + 
  guides(color=guide_legend(ncol=2)) + 
  theme(legend.position = 'top') + 
  xlab('sequence number') + 
  ylab(expression(delta^13*C[deviation])) + 
  ggtitle('Uncorrected')
```

```{r, echo = FALSE, message=FALSE}
# oxygen drift plot
O_drift_uncorrected <- data %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(aes(x = seq_nr, 
             y = d18O_dev,
             color = id2)) + 
  geom_point(size = 3) + 
  geom_smooth(method = 'lm', 
              formula = y ~ x, 
              se = FALSE,
              mapping = aes(group = 1),
              color = 'black') + 
  scale_color_manual(values = colors) +
  coord_cartesian(ylim = O_ylims) + 
  guides(color=guide_legend(ncol=2)) + 
  theme(legend.position = 'top') + 
  xlab('sequence number') + 
  ylab(expression(delta^18*O[deviation])) + 
  ggtitle('Uncorrected')
```

```{r, echo = FALSE, message = FALSE}

if(carbon_drift) {
  if(carbon_drift_model == 'lm') {
    C_drift_fit <- data %>% 
      filter(id2 %in% drift_material) %>%
      lm(d13C_dev ~ seq_nr, data = .)
  } else if(carbon_drift_model == 'loess') {
    C_drift_fit <- data %>% 
      filter(id2 %in% drift_material) %>%
      loess(d13C_dev ~ seq_nr, data = .)
  }
  data <- data %>% 
    mutate(d13C_measured = d13C_measured - predict(C_drift_fit, 
                                                   newdata = data)) %>%
    group_by(id2) %>%
    mutate(d13C_dev = d13C_measured - mean(d13C_measured)) %>%
    ungroup()
}

if(oxygen_drift) {
  if(oxygen_drift_model == 'lm') {
    O_drift_fit <- data %>% 
      filter(id2 %in% drift_material) %>%
      lm(d18O_dev ~ seq_nr, data = .)
  } else if(carbon_drift_model == 'loess') {
    O_drift_fit <- data %>% 
      filter(id2 %in% drift_material) %>%
      loess(d18O_dev ~ seq_nr, data = .)
  }
  data <- data %>% 
    mutate(d18O_measured = d18O_measured - predict(O_drift_fit, 
                                                   newdata = data)) %>%
    group_by(id2) %>%
    mutate(d18O_dev = d18O_measured - mean(d18O_measured)) %>%
    ungroup()
}
```

```{r, echo = FALSE, message=FALSE}
# Carbon drift plot
C_drift_corrected <- data %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(aes(x = seq_nr, 
             y = d13C_dev,
             color = id2)) + 
  geom_point(size = 3) + 
  geom_smooth(method = 'lm', 
              formula = y ~ x, 
              se = FALSE,
              mapping = aes(group = 1),
              color = 'black') + 
  scale_color_manual(values = colors) + 
  coord_cartesian(ylim = C_ylims) + 
  guides(color=guide_legend(ncol=2)) + 
  theme(legend.position = 'top') + 
  xlab('sequence number') + 
  ylab(expression(delta^13*C[deviation])) + 
  ggtitle('Corrected')
```

```{r, echo = FALSE, message=FALSE}
# oxygen drift plot
O_drift_corrected <- data %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(aes(x = seq_nr, 
             y = d18O_dev,
             color = id2)) + 
  geom_point(size = 3) + 
  geom_smooth(method = 'lm', 
              formula = y ~ x, 
              se = FALSE,
              mapping = aes(group = 1),
              color = 'black') + 
  scale_color_manual(values = colors) +
  coord_cartesian(ylim = O_ylims) + 
  guides(color=guide_legend(ncol=2)) + 
  theme(legend.position = 'top') + 
  xlab('sequence number') + 
  ylab(expression(delta^18*O[deviation])) + 
  ggtitle('Corrected')
```

## Carbon Drift Correction

Carbon isotope composition drift over the period of analysis was monitored using the **`r paste(drift_material, collapse = ' and ')`** reference material(s). Since a drift correction was `r ifelse(carbon_drift,' ','not')` requested, A **`r if(carbon_drift_model == 'lm') {paste('linear')} else if(carbon_drift_model == 'loess') {paste('loess')}`** model was `r ifelse(carbon_drift,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
if(carbon_drift){
  plot_grid(C_drift_uncorrected, C_drift_corrected)
}
```

## Oxygen Drift Correction

Oxygen isotope composition drift over the period of analysis was monitored using the **`r paste(drift_material, collapse = ' and ')`** reference material(s). Since a drift correction was `r ifelse(oxygen_drift,' ','not')` requested, A **`r if(oxygen_drift_model == 'lm') {paste('linear')} else if(oxygen_drift_model == 'loess') {paste('loess')}`** model was `r ifelse(oxygen_drift,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
if(oxygen_drift){
  plot_grid(O_drift_uncorrected, O_drift_corrected)
}
```

# Standardization {.tabset}

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
C_std_plot <- data %>% 
  filter(id2 %in% standards_used & is.finite(d13C_true)) %>% 
  ggplot(mapping = aes(x = d13C_measured, 
                       y = d13C_true,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors) + 
  geom_smooth(method = 'lm', 
              aes(group = 1), 
              color = 'black') + 
  theme(legend.position = 'top') + 
  guides(color=guide_legend(ncol=2)) + 
  xlab(expression(delta^13*C[measured])) + 
  ylab(expression(delta^13*C[accepted]))

O_std_plot <- data %>% 
  filter(id2 %in% standards_used & is.finite(d18O_true)) %>% 
  ggplot(mapping = aes(x = d18O_measured, 
                       y = d18O_true,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors) + 
  geom_smooth(method = 'lm', 
              aes(group = 1), 
              color = 'black') + 
  theme(legend.position = 'top') + 
  guides(color=guide_legend(ncol=2)) + 
  xlab(expression(delta^18*O[measured])) + 
  ylab(expression(delta^18*O[accepted]))

plot_grid(C_std_plot, O_std_plot)
rm(C_std_plot, O_std_plot)
```

```{r, echo = FALSE, message = FALSE}
C_std_fit <- data %>% 
  filter(id2 %in% standards_used & is.finite(d13C_true)) %>% 
  lm(d13C_true ~ d13C_measured, data = .)

O_std_fit <- data %>% 
  filter(id2 %in% standards_used & is.finite(d18O_true)) %>% 
  lm(d18O_true ~ d18O_measured, data = .)

data <- data %>% 
  mutate(d13C_corrected = predict(C_std_fit, newdata = data),
         d18O_corrected = predict(O_std_fit, newdata = data))
```

The isotope composition of unknowns samples were corrected to the international PDB ($\delta^{13}C$, $\delta^{18}O$) scale using the **`r paste(standards_used, collapse = ' and ')`** standard reference materials. Isotope compositions for standard reference materials are shown in the **Table 1**.

```{r, echo = FALSE}
standards %>% 
  filter(id2 %in% standards_used) %>%
  select(id2, 
         d13C_true, 
         d18O_true) %>% 
  knitr::kable(col.names = c('standard', 
                             '$\\delta^{13}C$', 
                             '$\\delta^{18}O$'),
               caption = 'Table 1: Certified isotope compositions for standard reference materials')
```

# Final Data {.tabset}

## Summary

```{r, echo = FALSE, message=FALSE}
final_data <- data %>% 
  select(Identifier_1 = id1, 
         sample = id2,
         d13C_corrected,
         d13C_sd,
         d18O_corrected,
         d18O_sd)

final_data %>% 
  filter(sample %in% c(standards_used, check_material)) %>% 
  group_by(sample) %>% 
  summarise(d13C_mean = mean(d13C_corrected),
            d13C_sd = sd(d13C_corrected),
            d18O_mean = mean(d18O_corrected),
            d18O_sd = sd(d18O_corrected),
            n = n()) %>% 
  knitr::kable(digits = 2, 
               col.names = c('sample', 
                             '$\\delta^{13}C_{mean}$', 
                             '$\\delta^{13}C_{sd}$', 
                             '$\\delta^{18}O_{mean}$', 
                             '$\\delta^{18}O_{sd}$', 
                             'n'),
               caption = 'Summary statistics for standard 
              reference materials and in house reference materials.')
```

## Full Data Table

```{r, echo = FALSE}
final_data %>%
  arrange(sample) %>%
  mutate_if(is.numeric, round, 3) %>% 
  knitr::kable(col.names = c('Identfier 1', 
                             'sample',
                             '$\\delta^{13}C$',
                             '$\\delta^{13}C_{sd}$',
                             '$\\delta^{18}O$',
                             '$\\delta^{18}O_{sd}$'),
               caption = 'Corrected isotope compositions.')
```

```{r, echo = FALSE, message = FALSE}
final_data %>% arrange(sample) %>% write_csv(file = paste0('./GB_results/', file_name,'_corrected.csv'))
raw_data %>% write_csv(file = paste0('./GB_results/', file_name, '_raw.csv'))
file.copy(from = 'GB_tidy_reduction.Rmd', to = paste0('./GB_results/', file_name,'_code.Rmd'))
```
