---
set_title: 'TCEA Report'
author: "Robin B. Trayler"
date: "`r format(Sys.time(), '%B, %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{css, echo=FALSE}
.block-format {
background-color: lightgrey2;
border: 2px solid black;
font-weight: bold;
}

/* Whole document: */
body{
font-family: Helvetica;
font-size: 12pt;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The $\delta^{2}H$ and $\delta^{18}O$ values  of all samples were measured using a **ThermoFisher TC/EA** coupled with a **Delta V Plus Continuous Flow Isotope Ratio Mass Spectrometer** housed in the Stable Isotope Facility at the University of California, Merced. 

This report includes several files:

1) A `.html` or `.pdf` report summarizing all data corrections and instrument reproducibility
2) A `.csv` file with the corrected data
3) A `.csv` file with the uncorrected data 
4) A `.Rmd` (R Markdown) document containing the **R** code used to correct the data

If you would like the original Isodat 3.0 `.dxf` files please request by sending an email to stableisotopes@ucmerced.edu


```{r, message = FALSE, echo = FALSE}
library(tidyverse)
library(viridis)
library(cowplot)
library(lubridate)
theme_set(theme_bw()) # set ggplot theme
```

```{r, echo = FALSE, message = FALSE}
standards <- tribble(~id2,      ~d18O_true, ~d2H_true, ~O_percent, ~H_percent,
                     'USGS 80',   13.1,        NA,       15.29,       0,
                     'USGS 81',   35.4,        NA,       15.29,       0,
                     'IAEA 601',  23.3,        NA,       26.20 ,      4.91,
                     'CBS',        3.8,       -157,        NA,         NA,
                     'KHS',       20.3,      -35.3,        NA,         NA,
                     'EA Ag3PO4', 21.7,        NA,        15.29,       0,
                     'USGS 54',  17.79,       -150.4,       NA,         NA,
                     'USGS 55',  19.12,        -28.2,       NA,         NA,
                     'EA Spruce', 23.87,          NA,      40.18,        6.29)
# the +1 gets rid of yellow
colors <- viridis(n = length(standards$id2) + 1)
names(colors) <- c(standards$id2, TRUE)
```

# Setup {.tabset}
## File Path
```{r}
# file path to the isodat csv file 
file_path <- '~/Box Sync/Data Repository/TCEA/TCEA20230329.csv'
```

## Analysis Parameters
### Select Sample Peaks
```{r}
# Which peak is the H2 analyte? 
hydrogen_peak <- 3
# Which peak is the CO analyte?
oxygen_peak <- 4
# How many total peaks should there be?
total_peaks <- 6
```

### Select Correction Standards
```{r}
# standards used in run
# standards_used = c('USGS 54', 'USGS 55', 'EA Spruce')
standards_used = c('CBS', 'KHS')
# material to use for linearity corrections
linearity_material =  c('IAEA 601')
# material to use for drift corrections
drift_material = c('CBS', 'KHS')
# in house 'check' materials for run quality
check_material = c('IAEA 601')

weight_percent_material = c('IAEA 601')
```

## Drift and Linearity Model Specification
### Instrument Drift
```{r}
# should a drift correction be applied to oxygen
oxygen_drift <- TRUE
hydrogen_drift <- TRUE
```

### Mass Linearity
```{r}
# should a linearity correction be applied?
# mass linearity corrections use a linear model
oxygen_linearity <- TRUE
oxygen_drift_model = 'lm'
hydrogen_drift_model = 'lm'
hydrogen_linearity <- TRUE
```

```{r, echo = FALSE}
# set the document file name to the name of the data file read in
file_name = tools::file_path_sans_ext(basename(file_path))
current_year <- format(Sys.Date(), '%Y')

#Check if directories exists and create if they doesn't
if(!dir.exists(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year))) {
  dir.create(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year))
}

if(!dir.exists(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name))) {
  dir.create(path = paste0(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name)))
}
```
---
title: "`r paste(file_name,' Report')`"
---

# Data Screening {.tabset}
```{r, message = FALSE, echo = FALSE}
# read data
raw_data <- read_csv(file = file_path)

data <- raw_data %>% 
  select(seq_nr = Row,
         id1 = `Identifier 1`, 
         id2 = `Identifier 2`,
         peak_nr = `Peak Nr.`,
         weight = Amount,
         area28 = `Area 28`,
         area2  = `Area 2`,
         d2H_measured = `d 2H/1H`,
         d18O_measured = `d 18O/16O`) |> 
  left_join(standards, by = 'id2')
```

```{r, echo = FALSE}
# seperate out the data
oxygen <- data %>% 
  select(id1,
         id2,
         peak_nr,
         seq_nr,
         weight,
         area28,
         d18O_measured,
         O_percent) |> 
  filter(peak_nr == oxygen_peak) %>% 
  group_by(id2) %>% 
  mutate(d18O_dev = d18O_measured - mean(d18O_measured)) %>% 
  ungroup() |> 
  mutate(weight_oxygen = O_percent * weight / 100)


hydrogen <- data |>  
  select(id1,
         id2,
         peak_nr,
         seq_nr,
         weight,
         area2,
         d2H_measured,
         H_percent) |> 
  filter(peak_nr == hydrogen_peak) |>  
  group_by(id2) |>  
  mutate(d2H_dev = d2H_measured - mean(d2H_measured)) |>  
  ungroup() |> 
  mutate(weight_hydrogen = H_percent * weight / 100)
```

# Drift Correction {.tabset}

```{r, echo=FALSE}
# figure out axis limits ------------------------------------------------------
O_ylims <- oxygen %>% 
  filter(id2 %in% drift_material) %>% 
  select(d18O_dev) %>% 
  range() %>% pretty() %>% range()

H_ylims <- hydrogen %>% 
  filter(id2 %in% drift_material) %>% 
  select(d2H_dev) %>% 
  range() %>% pretty() %>% range()
```

```{r, echo = FALSE, message=FALSE}
# oxygen pre calibration plot -------------------------------------------------
O_drift_uncorrected <- oxygen %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(mapping = aes(x = seq_nr,
                       y = d18O_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) + 
  xlab('sequence number') + 
  ylab(expression(delta^18*O[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = O_ylims) + 
  guides(color=guide_legend(ncol=2))

# hydrogen pre calibration plot -----------------------------------------------
H_drift_uncorrected <- hydrogen %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(mapping = aes(x = seq_nr,
                       y = d2H_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) +
  xlab('sequence number') + 
  ylab(expression(delta^2*H[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = H_ylims) + 
  guides(color=guide_legend(ncol=2))
```

```{r, message = FALSE, echo = FALSE}
# fit a Drift Correction ------------------------------------------------------
if(oxygen_drift) {
  # generate a calibration
  O_drift_fit <- oxygen %>% 
    filter(id2 %in% drift_material) %>%
    lm(d18O_dev ~ seq_nr, data = .)
  
  # apply the calibration
  oxygen <- oxygen %>% 
    mutate(d18O_measured = d18O_measured - predict(O_drift_fit, 
                                                   newdata = oxygen)) %>%
    # recalculate deviations
    group_by(id2) %>%
    mutate(d18O_dev = d18O_measured - mean(d18O_measured)) %>%
    ungroup()
}


if(hydrogen_drift) {
  # generate a calibration
  H_drift_fit <- hydrogen %>% 
    filter(id2 %in% drift_material) %>%
    lm(d2H_dev ~ seq_nr, data = .)
  
  # apply the calibration
  hydrogen <- hydrogen %>% 
    mutate(d2H_measured = d2H_measured - predict(H_drift_fit, 
                                                 newdata = hydrogen)) %>%
    # recalculate deviations
    group_by(id2) %>%
    mutate(d2H_dev = d2H_measured - mean(d2H_measured)) %>%
    ungroup()
}

```


```{r, echo = FALSE, message=FALSE}
if(oxygen_drift){
  # post correction plot --------------------------------------------------------
  O_drift_corrected <- oxygen %>% 
    filter(id2 %in% drift_material) %>% 
    ggplot(mapping = aes(x = seq_nr,
                         y = d18O_dev,
                         color = id2)) + 
    geom_point(size = 3) + 
    theme(legend.position = 'top') + 
    geom_smooth(method = 'lm',
                se = FALSE, 
                color = 'black') + 
    scale_color_manual(values = colors, limits = force) + 
    xlab('sequence number') + 
    ylab(expression(delta^18*O[deviation])) + 
    ggtitle('Corrected') + 
    coord_cartesian(ylim = O_ylims) + 
    guides(color=guide_legend(ncol=2))
}

if(hydrogen_drift){
  # post correction plot --------------------------------------------------------
  H_drift_corrected <- hydrogen %>% 
    filter(id2 %in% drift_material) %>% 
    ggplot(mapping = aes(x = seq_nr,
                         y = d2H_dev,
                         color = id2)) + 
    geom_point(size = 3) + 
    theme(legend.position = 'top') + 
    geom_smooth(method = 'lm',
                se = FALSE, 
                color = 'black') + 
    scale_color_manual(values = colors, limits = force) + 
    xlab('sequence number') + 
    ylab(expression(delta^2*H[deviation])) + 
    ggtitle('Corrected') + 
    coord_cartesian(ylim = H_ylims) + 
    guides(color=guide_legend(ncol=2))
}

```

## Oxygen Drift Correction
Oxygen isotope compositional drift over the period of analysis was monitored using the **`r paste(drift_material, collapse = ' and ')`** reference material(s). Since a drift correction was `r ifelse(oxygen_drift,' ','not')` requested, A **`r if(oxygen_drift_model == 'lm') {paste('linear')} else if(oxygen_drift_model == 'loess') {paste('loess')}`** model was `r ifelse(oxygen_drift,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
if(oxygen_drift){
  plot_grid(O_drift_uncorrected, O_drift_corrected)
}
```

## Hydrogen Drift Correction
Hydrogen isotope compositional drift over the period of analysis was monitored using the **`r paste(drift_material, collapse = ' and ')`** reference material(s). Since a drift correction was `r ifelse(hydrogen_drift,' ','not')` requested, A **`r if(oxygen_drift_model == 'lm') {paste('linear')} else if(hydrogen_drift_model == 'loess') {paste('loess')}`** model was `r ifelse(hydrogen_drift,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
if(hydrogen_drift){
  plot_grid(H_drift_uncorrected, H_drift_corrected)
}
```


# Linearity Corrections {.tabset}
```{r, echo = FALSE}
O_ylims <- oxygen %>% 
  filter(id2 %in% linearity_material) %>% 
  select(d18O_dev) %>% 
  range() %>% pretty() %>% range()

H_ylims <- hydrogen %>% 
  filter(id2 %in% linearity_material) %>% 
  select(d2H_dev) %>% 
  range() %>% pretty() %>% range()
```

```{r, echo = FALSE}
# Oxygen
# pre calibration plot --------------------------------------------------------
O_lin_uncorrected <- oxygen %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area28,
                       y = d18O_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab(expression(delta^18*O[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = O_ylims) + 
  guides(color=guide_legend(ncol=2))

H_lin_uncorrected <- hydrogen %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area2,
                       y = d2H_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab(expression(delta^12*H[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = H_ylims) + 
  guides(color=guide_legend(ncol=2))
```

```{r, echo = FALSE, message = FALSE}
# Oxygen
# fit a linearity correction --------------------------------------------------
O_lin_fit <- oxygen %>% 
  filter(id2 %in% linearity_material) %>% 
  lm(formula = d18O_dev ~ area28, data = .)
O_lin_sum <- summary(O_lin_fit)

H_lin_fit <- hydrogen %>% 
  filter(id2 %in% linearity_material) %>% 
  lm(formula = d2H_dev ~ area2, data = .)
H_lin_sum <- summary(H_lin_fit)

if(oxygen_linearity) {
  oxygen <- oxygen %>% 
    mutate(d18O_measured = d18O_measured - predict(O_lin_fit, newdata = oxygen)) %>% 
    group_by(id2) %>% 
    mutate(d18O_dev = d18O_measured - mean(d18O_measured)) %>% 
    ungroup()
}

if(hydrogen_linearity) {
  hydrogen <- hydrogen %>% 
    mutate(d2H_measured = d2H_measured - predict(H_lin_fit, newdata = hydrogen)) %>% 
    group_by(id2) %>% 
    mutate(d2H_dev = d2H_measured - mean(d2H_measured)) %>% 
    ungroup()
}
```

```{r, echo = FALSE}
# post calibration plot -------------------------------------------------------
O_lin_corrected <- oxygen %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area28,
                       y = d18O_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab(expression(delta^18*O[deviation])) + 
  ggtitle('Corrected') + 
  coord_cartesian(ylim = O_ylims) + 
  guides(color=guide_legend(ncol=2))

H_lin_corrected <- hydrogen %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area2,
                       y = d2H_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab(expression(delta^2*H[deviation])) + 
  ggtitle('Corrected') + 
  coord_cartesian(ylim = H_ylims) + 
  guides(color=guide_legend(ncol=2))

```

## Oxygen Linearity Correction
Mass linearity of oxygen isotope compositions was monitored using the **`r paste(linearity_material, collapse = ' and ')`** reference material(s). Since a linearity correction was `r ifelse(oxygen_linearity,' ','not')` requested, A linear trend with an $r^2$ of **`r round(O_lin_sum$r.squared, 2)`** was `r ifelse(oxygen_linearity,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center'}
if(oxygen_linearity) {
  plot_grid(O_lin_uncorrected, O_lin_corrected)
}

```

## Hydrogen Linearity Correction
Mass linearity of hydrogen isotope compositions was monitored using the **`r paste(linearity_material, collapse = ' and ')`** reference material(s). Since a linearity correction was `r ifelse(oxygen_linearity,' ','not')` requested, A linear trend with an $r^2$ of **`r round(H_lin_sum$r.squared, 2)`** was `r ifelse(hydrogen_linearity,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center'}
if(hydrogen_linearity) {
  plot_grid(H_lin_uncorrected, H_lin_corrected)
}
```

# Standardization {.tabset}

```{r, echo = FALSE}
# add the standards 
oxygen <- oxygen %>% 
  left_join(standards %>% 
              filter(id2 %in% standards_used), by = 'id2')

hydrogen <- hydrogen %>% 
  left_join(standards %>% 
              filter(id2 %in% standards_used), by = 'id2')
```

```{r, echo = FALSE, message=FALSE, fig.align='center'}
# calibration plot
O_scale <- oxygen %>% 
  filter(id2 %in% standards_used) %>% 
  ggplot(mapping = aes(x = d18O_measured,
                       y = d18O_true,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors, limits = force) +
  geom_smooth(method = 'lm', aes(group = 1), color = 'black') + 
  theme(legend.position = 'top') + 
  xlab(expression(delta^18*O[measured])) + 
  ylab(expression(delta^18*O[accepted])) + 
  guides(color=guide_legend(ncol=2))

H_scale <- hydrogen %>% 
  filter(id2 %in% standards_used) %>% 
  ggplot(mapping = aes(x = d2H_measured,
                       y = d2H_true,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors, limits = force) +
  geom_smooth(method = 'lm', aes(group = 1), color = 'black') + 
  theme(legend.position = 'top') + 
  xlab(expression(delta^2*H[measured])) + 
  ylab(expression(delta^2*H[accepted])) + 
  guides(color=guide_legend(ncol=2))

plot_grid(O_scale, H_scale)
```

```{r, echo = FALSE}
O_std_fit <- oxygen %>%  
  filter(id2 %in% standards_used) %>% 
  lm(formula = d18O_true ~ d18O_measured, data = .)

oxygen <- oxygen %>% 
  mutate(d18O_corrected = predict(O_std_fit, newdata = oxygen))


H_std_fit <- hydrogen %>%  
  filter(id2 %in% standards_used) %>% 
  lm(formula = d2H_true ~ d2H_measured, data = .)

hydrogen <- hydrogen %>% 
  mutate(d2H_corrected = predict(H_std_fit, newdata = hydrogen))

```

The isotope composition of unknowns samples were corrected to the international SMOW ($\delta^{18}O$) scale using the **`r paste(standards_used, collapse = ' and ')`** standard reference materials. Isotope compositions for standard reference materials are shown in the **Table 1**.

```{r, echo = FALSE}
standards %>% 
  filter(id2 %in% standards_used) %>%
  select(id2, 
         d18O_true,
         d2H_true) %>% 
  knitr::kable(col.names = c('standard', 
                             '$\\delta^{18}O$',
                             '$\\delta^{2}H$'),
               caption = 'Table 1: Certified isotope compositions for standard reference materials')
```

# Elemental Content

```{r echo=FALSE, fig.align='center', message=FALSE}
O_weight <- oxygen %>% 
  filter(id2 %in% weight_percent_material) %>% 
  ggplot(mapping = aes(x = area28, 
                       y = weight_oxygen,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors, limits = force) + 
  theme(legend.position = 'top') + 
  geom_smooth(aes(group = 1), color = 'black', method = 'lm') + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab('oxygen content (mg)') + 
  guides(color=guide_legend(ncol=2))

H_weight <- hydrogen %>% 
  filter(id2 %in% weight_percent_material) %>% 
  ggplot(mapping = aes(x = area2, 
                       y = weight_hydrogen,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors, limits = force) + 
  theme(legend.position = 'top') + 
  geom_smooth(aes(group = 1), color = 'black', method = 'lm') + 
  xlab('mass 2 peak area (mV/s)') + 
  ylab('hydrogen content (mg)') + 
  guides(color=guide_legend(ncol=2))

plot_grid(O_weight, H_weight)
```

```{r, echo = FALSE}
O_wt_fit <- oxygen %>% 
  filter(id2 %in% weight_percent_material) %>% 
  lm(formula = weight_oxygen ~ area28, data = .)
oxygen <- oxygen %>% 
  mutate(O_wt_percent = predict(O_wt_fit, newdata = oxygen) / weight * 100)

H_wt_fit <- hydrogen %>% 
  filter(id2 %in% weight_percent_material) %>% 
  lm(formula = weight_hydrogen ~ area2, data = .)
hydrogen <- hydrogen %>% 
  mutate(H_wt_percent = predict(H_wt_fit, newdata = hydrogen) / weight * 100)
```

Total oxygen and hydrogen contents were calculated using the **`r paste(weight_percent_material, collapse = ' and ')`** reference materials. Certified carbon and nitrogen contents for these reference materials are shown in the **Table 2**.

```{r, echo = FALSE}
standards %>% 
  filter(id2 %in% weight_percent_material) %>%
  select(id2, 
         O_percent, 
         H_percent) %>% 
  knitr::kable(col.names = c('standard', 
                             '$O \\%$', 
                             '$H \\%$'),
               caption = 'Table 2: Certified elemental contents for standard reference materials')
```



# Final Data {.tabset}

```{r, echo = FALSE, message = FALSE}

oxygen <- oxygen |> 
  select(Identifier_1 = id1,
         sample = id2, 
         amount = weight, 
         area_28 = area28,
         d18O_corrected,
         O_wt_percent)

hydrogen <- hydrogen |> 
  select(Identifier_1 = id1,
         sample = id2, 
         amount = weight, 
         area_2 = area2,
         d2H_corrected,
         H_wt_percent)


final_data <- oxygen |> full_join(hydrogen) |> 
  select(Identifier_1,
         sample,
         amount,
         area_28,
         area_2,
         d18O_corrected,
         d2H_corrected,
         O_wt_percent,
         H_wt_percent)

```

## Summary 
```{r, echo = FALSE, message = FALSE}
final_data %>% 
  filter(sample %in% c(standards_used, check_material, drift_material, linearity_material)) %>% 
  group_by(sample) %>% 
  summarise(d2H_mean = mean(d2H_corrected),
            d2H_sd   = sd(d2H_corrected),
            d18O_mean = mean(d18O_corrected),
            d18O_sd = sd(d18O_corrected),
            H_wt_percent_mean = mean(H_wt_percent),
            H_wt_percent_sd   = sd(H_wt_percent),
            O_wt_percent_mean = mean(O_wt_percent),
            O_wt_percent_sd   = sd(O_wt_percent),
            
            n = n()) %>% 
  knitr::kable(digits = 2, 
               col.names = c('sample', 
                             '$\\delta^{2}H_{mean}$', 
                             '$\\delta^{2}H_{sd}$', 
                             '$\\delta^{18}O_{mean}$', 
                             '$\\delta^{18}O_{sd}$', 
                             '$H \\%_{mean}$',
                             '$H \\%_{sd}$',
                             '$O \\%_{mean}$',
                             '$O \\%_{sd}$',
                             'n'),
               caption = 'Summary statistics for standard 
              reference materials and in house reference materials.')
```

## Full Data Table
```{r, echo = FALSE}
final_data %>%
  arrange(sample) %>%
  mutate_if(is.numeric, round, 3) %>% 
  knitr::kable(col.names = c('Identfier 1', 
                             'sample',
                             'amount (mg)', 
                             'area 28',
                             'area 2',
                             '$\\delta^{18}O$',
                             '$\\delta^{2}H$',
                             '$O \\%$',
                             '$H \\%$'), 
               caption = 'Corrected isotope compositions.')
```

```{r, echo = FALSE, message = FALSE}

final_data %>%
  arrange(sample) %>%
  write_csv(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name,'/',file_name,'_corrected.csv'))

raw_data %>%
  write_csv(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name,'/',file_name,'_raw.csv'))


file.copy(from = 'TCEA_tidy_reduction.Rmd',
          to = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name,'/',file_name,'_code.Rmd'))
```