---
set_title: 'TCEA Report'
author: "Robin B. Trayler"
date: "`r format(Sys.time(), '%B, %d, %Y')`"
output:
  html_document:
    # theme: cosmo
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{css, echo=FALSE}
.block-format {
  background-color: lightgrey2;
  border: 2px solid black;
  font-weight: bold;
}

/* Whole document: */
  body{
    font-family: Helvetica;
    font-size: 12pt;
  }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
The $\delta^{18}O$ values  of all samples were measured using a **ThermoFisher TC/EA** coupled with a **Delta V Plus Continuous Flow Isotope Ratio Mass Spectrometer** housed in the Stable Isotope Facility at the University of California, Merced. 

This report includes several files:

1) A `.html` or `.pdf` report summarizing all data corrections and instrument reproducibility
2) A `.csv` file with the corrected data
3) A `.csv` file with the uncorrected data 
4) A `.Rmd` (R Markdown) document containing the **R** code used to correct the data

If you would like the original Isodat 3.0 `.dxf` files please request by sending an email to stableisotopes@ucmerced.edu


```{r, message = FALSE, echo = FALSE}
library(tidyverse)
library(viridis)
library(cowplot)
library(lubridate)
theme_set(theme_bw()) # set ggplot theme
```

```{r, echo = FALSE}
z_score <- function(x) {
  abs(x - mean(x)) / sd(x)
}
```

```{r, echo = FALSE, message = FALSE}
# standards <- read_csv(file = './data/EA_standards.csv')
standards <- tribble(~id2,      ~d18O_true,
                     'USGS 80',   13.1, 
                     'USGS 81',   35.4,
                     'IAEA 601',  23.3,
                     'CBS',       3.8,
                     'KHS',       20.3,
                     'EA Ag3PO4', 21.7)

colors <- viridis(n = length(standards$id2) + 1)
names(colors) <- c(standards$id2, TRUE)
```

# Setup {.tabset}
## File Path
```{r}
# file path to the isodat csv file 
file_path <- '~/Box Sync/Data Repository/TCEA/TCEA20230317.csv'
```

## Analysis Parameters
### Select Sample Peaks
```{r}
# Which peak is the CO analyte?
oxygen_peak <- 3
# How many total peaks should there be?
total_peaks <- 5
```

### Set Rejection Criteria
```{r}
# rejection criteria
small_peaks <- 500 # warn about peaks smaller with an amplitude smaller than this
# should 2 sigma outlier standards be dropped from isotope corrections
drop_outliers <- FALSE 
```

### Select Correction Standards
```{r}
# standards used in run
standards_used = c('USGS 80', 'USGS 81')
# material to use for linearity corrections
linearity_material =  c('USGS 81', 'USGS 80')
# material to use for drift corrections
drift_material = c('USGS 81', 'USGS 80')
# in house 'check' materials for run quality
check_material = c('EA Ag3PO4', 'NIST 120C', 'alfa asar', 'IAEA 601')
```

## Drift and Linearity Model Specification
### Instrument Drift
```{r}
# should a drift correction be applied to oxygen
oxygen_drift <- TRUE
# model choices are either 'lm' for linear or 'loess' for loess
oxygen_drift_model <- 'lm'
```

### Mass Linearity
```{r}
# should a linearity correction be applied?
# mass linearity corrections use a linear model
oxygen_linearity <- TRUE
```

```{r, echo = FALSE}
# set the document file name to the name of the data file read in
file_name = tools::file_path_sans_ext(basename(file_path))
current_year <- format(Sys.Date(), '%Y')

#Check if directories exists and create if they doesn't
if(!dir.exists(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year))) {
  dir.create(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year))
}

if(!dir.exists(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name))) {
  dir.create(path = paste0(path = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name)))
}
```
---
title: "`r paste(file_name,' Report')`"
---

# Data Screening {.tabset}
```{r, message = FALSE, echo = FALSE}
# read data
raw_data <- read_csv(file = file_path)

data <- raw_data %>% 
  select(seq_nr = Row,
         id1 = `Identifier 1`, 
         id2 = `Identifier 2`,
         peak_nr = `Peak Nr.`,
         weight = Amount,
         area28 = `Area 28`,
         amp28 = `Ampl. 28`,
         amp30 = `Ampl. 30`,
         area30 = `Area 30`,
         d18O_measured = `d 18O/16O`)
```

## Check for Missing Peaks
```{r, echo = FALSE, message = FALSE}
# check for missing peaks
missing <- data %>% 
  group_by(id1) %>% 
  summarize(n = !(n() == total_peaks))
```
Sample(s) with an Identifier 1 of __`r missing$id1[missing$n]`__ have fewer peaks than expected. Please check the raw data.

## Check for Small Peaks
```{r, echo = FALSE}
small_O <- data %>% 
  filter(peak_nr %in% oxygen_peak & amp28 < small_peaks) %>% 
  select(id1)
```

Sample(s) with an Identifier 1 of __`r small_O$id1`__ have a mass 28 (carbon monoxide) amplitude smaller than __`r small_peaks`__ mV. Data from these samples may be unreliable.

# Drift Correction {.tabset}

```{r, echo = FALSE}
data <- data %>% 
  filter(peak_nr == oxygen_peak) %>% 
  group_by(id2) %>% 
  mutate(d18O_dev = d18O_measured - mean(d18O_measured)) %>% 
  ungroup()
```


```{r, echo=FALSE}
# figure out axis limits ------------------------------------------------------
O_ylims <- data %>% 
  filter(id2 %in% drift_material) %>% 
  select(d18O_dev) %>% 
  range() %>% pretty() %>% range()
```

```{r, echo = FALSE, message=FALSE}
# oxygen pre calibration plot -------------------------------------------------
O_drift_uncorrected <- data %>% 
  filter(id2 %in% drift_material) %>% 
  ggplot(mapping = aes(x = seq_nr,
                       y = d18O_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = oxygen_drift_model,
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) + 
  xlab('sequence number') + 
  ylab(expression(delta^18*O[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = O_ylims) + 
  guides(color=guide_legend(ncol=2))
```

```{r, message = FALSE, echo = FALSE}
# fit a Drift Correction ------------------------------------------------------
if(oxygen_drift) {
  if(oxygen_drift_model == 'lm') {
    O_drift_fit <- data %>% 
      filter(id2 %in% drift_material) %>%
      lm(d18O_dev ~ seq_nr, data = .)
  } else if(oxygen_drift_model == 'loess') {
    O_drift_fit <- data %>% 
      filter(id2 %in% drift_material) %>%
      loess(d18O_dev ~ seq_nr, data = .)
  }
  data <- data %>% 
    mutate(d18O_measured = d18O_measured - predict(O_drift_fit, newdata = data)) %>%
    group_by(id2) %>%
    mutate(d18O_dev = d18O_measured - mean(d18O_measured)) %>%
    ungroup()
}
```


```{r, echo = FALSE, message=FALSE}
if(oxygen_drift){
  # post correction plot --------------------------------------------------------
  O_drift_corrected <- data %>% 
    filter(id2 %in% drift_material) %>% 
    ggplot(mapping = aes(x = seq_nr,
                         y = d18O_dev,
                         color = id2)) + 
    geom_point(size = 3) + 
    theme(legend.position = 'top') + 
    geom_smooth(method = 'lm',
                se = FALSE, 
                color = 'black') + 
    scale_color_manual(values = colors, limits = force) + 
    xlab('sequence number') + 
    ylab(expression(delta^18*O[deviation])) + 
    ggtitle('Corrected') + 
    coord_cartesian(ylim = O_ylims) + 
    guides(color=guide_legend(ncol=2))
}

```

## Oxygen Drift Correction
Oxygen isotope compositional drift over the period of analysis was monitored using the **`r paste(drift_material, collapse = ' and ')`** reference material(s). Since a drift correction was `r ifelse(oxygen_drift,' ','not')` requested, A **`r if(oxygen_drift_model == 'lm') {paste('linear')} else if(oxygen_drift_model == 'loess') {paste('loess')}`** model was `r ifelse(oxygen_drift,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center', warning = FALSE}
if(oxygen_drift){
  plot_grid(O_drift_uncorrected, O_drift_corrected)
}
```

# Linearity Corrections {.tabset}
```{r, echo = FALSE}
O_ylims <- data %>% 
  filter(id2 %in% linearity_material) %>% 
  select(d18O_dev) %>% 
  range() %>% pretty() %>% range()
```

```{r, echo = FALSE}
# Oxygen
# pre calibration plot --------------------------------------------------------
O_lin_uncorrected <- data %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area28,
                       y = d18O_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab(expression(delta^18*O[deviation])) + 
  ggtitle('Uncorrected') + 
  coord_cartesian(ylim = O_ylims) + 
  guides(color=guide_legend(ncol=2))
```

```{r, echo = FALSE, message = FALSE}
# Oxygen
# fit a linearity correction --------------------------------------------------
O_lin_fit <- data %>% 
  filter(id2 %in% linearity_material) %>% 
  lm(formula = d18O_dev ~ area28, data = .)
O_lin_sum <- summary(O_lin_fit)

if(oxygen_linearity) {
  data <- data %>% 
    mutate(d18O_measured = d18O_measured - predict(O_lin_fit, newdata = data)) %>% 
    group_by(id2) %>% 
    mutate(d18O_dev = d18O_measured - mean(d18O_measured)) %>% 
    ungroup()
}
```

```{r, echo = FALSE}
# post calibration plot -------------------------------------------------------
O_lin_corrected <- data %>% 
  filter(id2 %in% linearity_material) %>% 
  ggplot(mapping = aes(x = area28,
                       y = d18O_dev,
                       color = id2)) + 
  geom_point(size = 3) + 
  theme(legend.position = 'top') + 
  geom_smooth(method = 'lm',
              se = FALSE, 
              color = 'black') + 
  scale_color_manual(values = colors, limits = force) + 
  xlab('mass 28 peak area (mV/s)') + 
  ylab(expression(delta^18*O[deviation])) + 
  ggtitle('Corrected') + 
  coord_cartesian(ylim = O_ylims) + 
  guides(color=guide_legend(ncol=2))

```

## Oxygen Linearity Correction
Mass linearity of oxygen isotope compositions  was monitored using the **`r paste(linearity_material, collapse = ' and ')`** reference material(s). Since a linearity correction was `r ifelse(oxygen_linearity,' ','not')` requested, A linear trend with an $r^2$ of **`r round(O_lin_sum$r.squared, 2)`** was `r ifelse(oxygen_linearity,' ','not')` subtracted from the raw data.

```{r, echo = FALSE, message = FALSE, fig.align='center'}
if(oxygen_linearity) {
  plot_grid(O_lin_uncorrected, O_lin_corrected)
}
```

# Standardization {.tabset}

```{r, echo = FALSE}
# add the standards 
data <- data %>% full_join(standards %>% filter(id2 %in% standards_used), by = 'id2')
```

```{r, echo = FALSE, message=FALSE, fig.align='center'}
# calibration plot
O_scale <- data %>% 
  filter(id2 %in% standards_used) %>% 
  ggplot(mapping = aes(x = d18O_measured,
                       y = d18O_true,
                       color = id2)) + 
  geom_point(size = 3) + 
  scale_color_manual(values = colors, limits = force) +
  geom_smooth(method = 'lm', aes(group = 1), color = 'black') + 
  theme(legend.position = 'top') + 
  xlab(expression(delta^18*O[measured])) + 
  ylab(expression(delta^18*O[accepted])) + 
  guides(color=guide_legend(ncol=2)) + 
  coord_fixed()

plot_grid(O_scale)
```

```{r, echo = FALSE}
if(drop_outliers){
  O_std_fit <- data %>%  
    filter(id2 %in% standards_used) %>% 
    group_by(id2) %>% 
    mutate(z_score = z_score(d18O_measured)) %>% 
    filter(z_score < 2) %>% 
    ungroup() %>% 
    lm(formula = d18O_true ~ d18O_measured, data = .)
  
  data <- data %>%
    mutate(d18O_corrected = predict(O_std_fit, newdata = data))
  
} else {
  O_std_fit <- data %>%  
    filter(id2 %in% standards_used) %>% 
    lm(formula = d18O_true ~ d18O_measured, data = .)
  
  data <- data %>% 
    mutate(d18O_corrected = predict(O_std_fit, newdata = data))
  
}

```

```{r, echo = FALSE}
O_r2 <-O_std_fit %>% summary() %>% pluck('r.squared')
```

The isotope composition of unknowns samples were corrected to the international SMOW ($\delta^{18}O$) scale using the **`r paste(standards_used, collapse = ' and ')`** standard reference materials. Isotope compositions for standard reference materials are shown in the **Table 1**.

```{r, echo = FALSE}
standards %>% 
  filter(id2 %in% standards_used) %>%
  select(id2, 
         d18O_true) %>% 
  knitr::kable(col.names = c('standard', 
                             '$\\delta^{18}O$'),
               caption = 'Table 1: Certified isotope compositions for standard reference materials')
```

# Final Data {.tabset}
```{r, echo = FALSE, message = FALSE}

final_data <- data %>% 
  select(Identifier_1 = id1, 
         sample = id2, 
         amount = weight,
         `Area 28` = area28,
         `Area 30`= area30,
         d18O_corrected)
```

## Summary 
```{r, echo = FALSE, message = FALSE}
final_data %>% 
  filter(sample %in% c(standards_used, check_material, drift_material, linearity_material)) %>% 
  group_by(sample) %>% 
  summarise(d18O_mean = mean(d18O_corrected),
            d18O_sd = sd(d18O_corrected),
            n = n()) %>% 
  knitr::kable(digits = 2, 
               col.names = c('sample', 
                             '$\\delta^{18}O_{mean}$', 
                             '$\\delta^{18}O_{sd}$', 
                             'n'),
               caption = 'Summary statistics for standard 
              reference materials and in house reference materials.')
```

## Full Data Table
```{r, echo = FALSE}
final_data %>%
  arrange(sample) %>%
  mutate_if(is.numeric, round, 3) %>% 
  knitr::kable(col.names = c('Identfier 1', 
                             'sample',
                             'amount (mg)', 
                             'area 28',
                             'area 30',
                             '$\\delta^{18}O$'), 
               caption = 'Corrected isotope compositions.')
```

```{r, echo = FALSE, message = FALSE}

final_data %>% 
  arrange(sample) %>% 
  write_csv(file = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name,'/',file_name,'_corrected.csv'))

raw_data %>% 
  write_csv(file = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name,'/',file_name,'_raw.csv'))


file.copy(from = 'TCEA_tidy_reduction.Rmd', 
          to = paste0('~/Box Sync/Data Repository/TCEA/Corrected/',current_year,'/',file_name,'/',file_name,'_code.Rmd'))
```